SunPy - Python for Solar Physics

### 2014-08-28 2-1-Map.tex
edits from full read thru

\subsection{Map}\label{ssec:map}
The map data type stores 2D spatial data, such as images of the Sun and 
inner heliosphere. It provides: a wrapper around a \texttt{numpy} data array, 
the images associated spatial coordinates, and other metadata. The \texttt{Map} 
class provides methods for typical operations on 2D data, such as rotation and 
re-sampling, as well as visualisation functionality.
The \texttt{Map} class also provides a convenient interface for loading data 
from a variety of sources, including from FITS (Flexible Image Transport System) files, the standard format for storing image data in solar physics and astrophysics. An example of creating a \texttt{Map} object from a FITS file is shown in 
Listing~\ref{code:aia_1}.

The architecture of the map subpackage consists of a template map called
\texttt{GenericMap}, which is a subclass of \texttt{astropy.nddata.NDData}. 
\texttt{NDData} is a generic wrapper around a \texttt{numpy.ndarray} with a 
\texttt{meta} attribute to store metadata.
As \texttt{NDData} is currently still in development, \texttt{GenericMap} does 
not yet make full use of its capabilities, but this inheritance structure 
provides for future integration with \texttt{astropy}. In order to provide 
instrument- or detector-specific integration, \texttt{GenericMap} is designed
to be subclassed. Each subclass of \texttt{GenericMap} can register 
with the \texttt{Map} creation factory, which will then automatically return an instance
of the specific \texttt{GenericMap} subclass dependent upon the data provided. 
SunPy v0.4 has \texttt{GenericMap} specialisations for the following 
instruments:

\textit{Yohkoh}/SXT\footnote{SXT: Solar X-ray Telescope \citep{1991SoPh..136....1O, 1991SoPh..136...37T}.},
\textit{SOHO}/EIT and LASCO\footnote{SOHO: Solar and Heliospheric Observatory \citep{domingo1995}, EIT: Extreme Ultraviolet Telescope \citep{1995SoPh..162..291D} and LASCO: Large Angle Spectroscopic COronagraph \citep{1995SoPh..162..357B}.},
\textit{RHESSI}\footnote{RHESSI: Reuven Ramaty High Energy Solar Spectroscopic Image \citep{2002SoPh..210....3L}.},
\textit{STEREO}/EUVI and COR1/2\footnote{STEREO: Solar TErrestrial RElations Observatory \citep{2005AdSpR..36.1483K}, EUVI: Extreme Ultraviolet Imager \citep{2004SPIE.5171..111W} and COR1/2: CORonagraph 1/2 \citep{2002AdSpR..29.2017H}.},
\textit{Hinode}/XRT\footnote{XRT: X-Ray Telescope \citep{2007SoPh..243....3K, 2007SoPh..243...63G}.},
\textit{PROBA2}/SWAP\footnote{PROBA2: PRojects for On Board Autonomy 2 \citep{2013SoPh..286....5S}, SWAP: Sun Watcher Active Pixel \citep{2013SoPh..286...43S}.},
\textit{SDO}/AIA and HMI\footnote{HMI: Helioseismic Magnetic Imager, \citep{2012SoPh..275..207S}.}, and
\textit{IRIS}\footnote{IRIS: Interface Region Imaging Spectrograph \citep{2011SPD....42.1512L}.} SJI (slit-jaw imager) frames.

@@

\begin{listing}[H]
\pythoncode{pycode_map1.txt}
\begin{center}
\includegraphics[width=0.8\columnwidth]{aia_map_example}
\end{center}
\caption{Example of the \texttt{AIAMap} specialisation of 
\texttt{GenericMap}. First, a map is created from an \textit{SDO}/AIA FITS file. In this case, a demonstration file contained within the SunPy repository is used. A cutout
of the full map is then created by specifying the desired solar-$x$ and solar-$y$ ranges of the plot in data coordinates (in this case, arcseconds), and then a quick-view plot is created with lines of heliographic longitude and latitude over-plotted.}
\label{code:aia_1}
\end{listing}

>>>>

\subsection{Map}\label{ssec:map}
The map data type stores 2D spatial data, such as images of the Sun and 
inner heliosphere. It provides: a wrapper around a \texttt{numpy} data array, 
the images associated spatial coordinates, and other metadata. The \texttt{Map} 
class provides methods for typical operations on 2D data, such as rotation and 
re-sampling, as well as visualisation.
The \texttt{Map} class also provides a convenient interface for loading data 
from a variety of sources, including from FITS (Flexible Image Transport System) 
files, the standard format for storing image data in solar physics and astrophysics community. 
An example of creating a \texttt{Map} object from a FITS file is shown in 
Listing~\ref{code:aia_1}.

The architecture of the map subpackage consists of a template map called
\texttt{GenericMap}, which is a subclass of \texttt{astropy.nddata.NDData}. 
\texttt{NDData} is a generic wrapper around a \texttt{numpy.ndarray} with a 
\texttt{meta} attribute to store metadata.
As \texttt{NDData} is currently still in development, \texttt{GenericMap} does 
not yet make full use of its capabilities, but this inheritance structure 
provides for future integration with \texttt{astropy}. In order to provide 
instrument- or detector-specific integration, \texttt{GenericMap} is designed
to be subclassed. Each subclass of \texttt{GenericMap} can register 
with the \texttt{Map} creation function, which will then automatically return an instance
of the specific \texttt{GenericMap} subclass dependent upon the data provided. 
SunPy v0.5 has \texttt{GenericMap} specialisations for the following 
instruments:

\begin{itemize}
\item \textit{Yohkoh}/SXT - SXT: Solar X-ray Telescope \citep{1991SoPh..136....1O, 1991SoPh..136...37T},
\item \textit{SOHO}/EIT and LASCO - SOHO: Solar and Heliospheric Observatory \citep{domingo1995}, EIT: Extreme Ultraviolet Telescope \citep{1995SoPh..162..291D} and LASCO: Large Angle Spectroscopic COronagraph \citep{1995SoPh..162..357B}
\item \textit{RHESSI} - RHESSI: Reuven Ramaty High Energy Solar Spectroscopic Image \citep{2002SoPh..210....3L}.},
\item \textit{STEREO}/EUVI and COR1/2 - STEREO: Solar TErrestrial RElations Observatory \citep{2005AdSpR..36.1483K}, EUVI: Extreme Ultraviolet Imager \citep{2004SPIE.5171..111W} and COR1/2: CORonagraph 1/2 \citep{2002AdSpR..29.2017H}
\item \textit{Hinode}/XRT - XRT: X-Ray Telescope \citep{2007SoPh..243....3K, 2007SoPh..243...63G}.
\item \textit{PROBA2}/SWAP - PROBA2: PRojects for On Board Autonomy 2 \citep{2013SoPh..286....5S}, SWAP: Sun Watcher Active Pixel \citep{2013SoPh..286...43S}
\item \textit{SDO}/AIA and HMI - HMI: Helioseismic Magnetic Imager, \citep{2012SoPh..275..207S}
\item \textit{IRIS} - IRIS: Interface Region Imaging Spectrograph \citep{2011SPD....42.1512L} SJI (slit-jaw imager) frames.           
\end{itemize}

@@

\begin{listing}[H]
\pythoncode{pycode_map1.txt}
\begin{center}
\includegraphics[width=0.8\columnwidth]{aia_map_example}
\end{center}
\caption{Example of the \texttt{AIAMap} specialisation of 
\texttt{GenericMap}. First, a map is created from a sample \textit{SDO}/AIA FITS file. In this case, a demonstration file contained within the SunPy repository is used. A cutout
of the full map is then created by specifying the desired solar-$x$ and solar-$y$ ranges of the plot in data coordinates (in this case, arcseconds), and then a quick-view plot is created with lines of heliographic longitude and latitude over-plotted.}
\label{code:aia_1}
\end{listing}

### 2014-08-28 2-3-Spectra.tex
created code text files
spectra data source references
edits from full read thru

As with other SunPy data types, the \texttt{Spectrogram} class has been
built so that each instrument initialises using a subclass containing the instrument-specific 
functionalities. The common functionality provided by the base \texttt{Spectrogram} class includes
joining different time ranges and frequencies, performing frequency-dependent background subtraction,
and convenient visualization and sampling of the data.
Currently, the \texttt{Spectrogram} class supports radio spectrograms from the 
\href{http://www.e-callisto.org/}{e-Callisto}
solar radio spectrometer network and \textit{STEREO}/SWAVES spectrograms.

Listing \ref{code:spectra} shows how the \texttt{CallistoSpectrogram}
object retrieves spectrogram data in the time range specified taken at
the observatory of interest.  When the data is requested using the
\texttt{from\_range()} function, the object merges all the downloaded
files into a single spectrogram, across time and frequency.
In the example shown, data is provided in two frequency ranges:
20--90\,MHz and 55--355\,MHz.  Since the data are not evenly spaced in
the frequency range, the \texttt{Spectrogram} object linearises the
frequency axis to assist analysis.  The example also demonstrates
the implemented background subtraction method, which calculates
a constant background over time for each frequency channel.
% RJH: Which subtraction method?  If the specifics are irrelevant, cut it.
% RJH: is this better?
%ji to DPS - what is the linearisation that is happening here?
%ji- the linearisation consist in the frequency axis being linear and the 
% data being streched with it.  The frequency normally are not evenly spaced
%ays: I deleted the phrase "based on the values closer to the average" because I don't understand what that means

\begin{listing}[H]
\pythoncode{spectra.py}
\begin{center}
\includegraphics[width=0.8\columnwidth]{callisto_nobg}
\end{center}
\caption{Example of how \texttt{CallistoSpectrogram} retrieves the
  data for the requested time range and observatory, merges it, and
  removes the background signal.  The data requested -- `BIR' -- is
  the code name of the \href{http://www.rosseobservatory.ie}{Rosse Observatory}
  at Birr Castle in Ireland.}
\label{code:spectra}
\end{listing}

>>>>

As with other SunPy data types, the \texttt{Spectrogram} class has been
built so that each instrument initialises using a subclass containing the instrument-specific 
functionalities. The common functionality provided by the base \texttt{Spectrogram} class includes
joining different time ranges and frequencies, performing frequency-dependent background subtraction,
and convenient visualization and sampling of the data.
Currently, the \texttt{Spectrogram} class supports radio spectrograms from the 
\href{http://www.e-callisto.org/}{e-Callisto}
solar radio spectrometer network \citep{2009EM&P..104..277B} and \textit{STEREO}/SWAVES spectrograms \citep{2008SSRv..136..487B}.

Listing \ref{code:spectra} shows how the \texttt{CallistoSpectrogram}
object retrieves spectrogram data in the time range specified.  
When the data is requested using the
\texttt{from\_range()} function, the object merges all the downloaded
files into a single spectrogram, across time and frequency.
In the example shown, data is provided in two frequency ranges:
20--90\,MHz and 55--355\,MHz.  Since the data are not evenly spaced in
the frequency range, the \texttt{Spectrogram} object linearises the
frequency axis to assist analysis.  The example also demonstrates
the implemented background subtraction method, which calculates
a constant background over time for each frequency channel.

\begin{listing}[H]
\pythoncode{pycode_spectra.txt}
\begin{center}
\includegraphics[width=0.8\columnwidth]{callisto_nobg}
\end{center}
\caption{Example of how \texttt{CallistoSpectrogram} retrieves the
  data for the requested time range and observatory, merges it, and
  removes the background signal.  The data requested -- `BIR' -- is
  the code name of the \href{http://www.rosseobservatory.ie}{Rosse Observatory}
  at Birr Castle in Ireland.}
\label{code:spectra}
\end{listing}

### 2014-08-28 2-Data_Types.tex
edits from full read thru

The core of SunPy is a set of data structures that are specifically
designed for the three primary varieties of solar physics data:
images, time series, and spectra. These core data types are supported
by the SunPy classes: \texttt{Map} (2D spatial data),
\texttt{LightCurve} (1D temporal series), and \texttt{Spectrum} and
\texttt{Spectrogram} (1D and 2D spectra).  The purpose of these
classes is to provide the same core data type to the SunPy user
regardless of the differences in source data.  For example, if two
different instruments use different time formats to describe the
observation time of their images, the corresponding SunPy \texttt{Map}
object for each of them expresses the observation time in the same
way.  This simplifies the workflow for the user when handling data
from multiple sources.

These classes allow access to the data
and associated metadata and provide appropriate convenience functions to
enable analysis and visualisation. For each of these classes, the data is
stored in the \texttt{data} attribute, while the metadata should be stored 
in the \texttt{meta} attribute (currently, only \texttt{Map} has this feature 
fully implemented). 
% schriste - change this until all fixed
It is possible to instantiate the
data types from various
different sources: e.g., files, URLs, and arrays.  
In order to provide instrument-specific specialisation, the core SunPy classes 
make use of subclassing; e.g., \texttt{Map} has an \texttt{AIAMap} 
sub-type for data from the \textit{SDO}/AIA (Atmospheric Imaging Assembly; \citealt{lemen2012}) instrument.

All of the core SunPy data types 
include visualisation methods that are tailored to each data type. 
These visualisation methods all currently utilise the \texttt{matplotlib} 
package and are designed in such a way that they integrate well with 
the \texttt{pyplot} functional interface of \texttt{matplotlib}.

>>>>

The core of SunPy is a set of data structures that are specifically
designed for the three primary varieties of solar physics data:
images, time series, and spectra. These core data types are supported
by the SunPy classes: \texttt{Map} (2D spatial data),
\texttt{LightCurve} (1D temporal series), and \texttt{Spectrum} and
\texttt{Spectrogram} (1D and 2D spectra).  The purpose of these
classes is to provide the same core data type to the SunPy user
regardless of the differences in source data.  This simplifies 
the workflow for the user when handling data from multiple sources.

These classes allow access to the data
and associated metadata and provide appropriate convenience functions to
enable analysis and visualisation. For each of these classes, the data is
stored in the \texttt{data} attribute, while the metadata is stored 
in the \texttt{meta} attribute\footnote{Note, that currently only \texttt{Map} and \texttt{LightCurve} have this feature 
fully implemented}. 
It is possible to instantiate the
data types from various
different sources: e.g., files, URLs, and arrays.  
In order to provide instrument-specific specialisation, the core SunPy classes 
make use of subclassing; e.g., \texttt{Map} has an \texttt{AIAMap} 
sub-type for data from the \textit{SDO}/AIA (Atmospheric Imaging Assembly; \citealt{lemen2012}) instrument.

All of the core SunPy data types 
include visualisation methods that are tailored to each data type. 
These visualisation methods all utilise the \texttt{matplotlib} 
package and are designed in such a way that they integrate well with 
the \texttt{pyplot} functional interface of \texttt{matplotlib}.

### 2014-08-28 3-1-VSO.tex
Missing comma after e.g.
edits from full read thru

The Virtual Solar Observatory (\href{http://virtualsolar.org}{VSO}) provides a 
single, standard query interface to solar data from different archives 
around the world \citep{hill2009}.
Data products can be requested for specific instruments or missions and
can also be requested based on physical parameters of the data product such
as the wavelength range.
In addition to the VSO's primary web-based interface, a SOAP (Simple Object 
Access Protocol) service is also available.
SunPy's \texttt{vso} module provides access to the VSO via this SOAP service using the
\texttt{suds} package.

Listing~\ref{code:vso_query_simple} shows an example of how to query and download data using
from the VSO using the \texttt{vso} module.
Queries are constructed using one or more attribute objects. Each
attribute object is a constraint on a parameter of the data set, such as the
time, instrument, or wavelength.
Listing~\ref{code:vso_query_simple} also shows how to download the data using
the constructed query. The path to which the data files will be downloaded us defined using custom tokens
which reference the file metadata (e.g. instrument, detector, filename). This provides
users the ability to organize their data into subdirectories on download.

>>>>

The Virtual Solar Observatory (\url{http://virtualsolar.org}{VSO}) provides a 
single, standard query interface to solar data from many different archives 
around the world \citep{hill2009}.
Data products can be requested for specific instruments or missions and
can also be requested based on physical parameters of the data product such
as the wavelength range.
In addition to the VSO's primary web-based interface, a SOAP (Simple Object 
Access Protocol) service is also available.
SunPy's \texttt{vso} module provides access to the VSO via this SOAP service using the
\texttt{suds} package.

Listing~\ref{code:vso_query_simple} shows an example of how to query and download data using
from the VSO using the \texttt{vso} module.
Queries are constructed using one or more attribute objects. Each
attribute object is a constraint on a parameter of the data set, such as the
time, instrument, or wavelength.
Listing~\ref{code:vso_query_simple} also shows how to download the data using
the constructed query. The path to which the data files will be downloaded is defined using custom tokens
which reference the file metadata (e.g., instrument, detector, filename). This provides
users the ability to organize their data into subdirectories on download.

### 2014-08-28 3-2-HEK.tex
edits from full read thru

The Sun is an active star and exhibits a wide range of transient phenomena 
(e.g., flares, radio bursts, coronal mass ejections) at many different time-scales, 
length-scales, and 
wavelengths. Observations and metadata concerning these phenomena are collected 
in the Heliophysics Event Knowledgebase (HEK) \citep{hek2012}.  Entries are generated both by 
automated algorithms and human observers.  Some of the information in the HEK 
reproduces feature and event data from elsewhere (for example, the \textit{GOES} flare catalogue),
and some is generated by the Solar Dynamics Observatory Feature Finding Team 
\citep{martens2012}.  A key feature of the HEK is that it
provides an homogeneous and well-described interface to a large amount of 
feature and event information of interest to the solar physics community. SunPy 
accesses this information through the \texttt{hek} module.  The \texttt{hek} module makes use of the 
HEK public API\footnote{For more information see \url{http://vso.stanford.edu/hekwiki/ApplicationProgrammingInterface}}.

@@

HEK comparisons can be combined using Python's logical operators (e.g., \texttt{and}
and \texttt{or}). The ability to use comparison and logical operators on HEK attributes allows 
the construction of queries of almost arbitrary complexity.
For the query in Listing \ref{code:hek:or} returns
returns flares with helioprojective x coordinates west of 50 arcseconds or 
those that have a peak flux above 1000.0 in units defined by the FRM.

@@

Frequently after users have found events of interest the next step is to 
download observational data. The \texttt{H2VClient} module makes this
much easier by providing a translation layer between HEK query results
and VSO queries. This capability is demonstrated in Listing~\ref{code:hek2vso}.
\begin{listing}[H]
\pythoncode{pycode_hek4.txt}
\caption{Code snippet continuing from Listing~\ref{code:hek:or} showing the 
query and download of data from the first HEK result from the VSO.}
\label{code:hek2vso}
\end{listing}

>>>>

The Sun is an active star and exhibits a wide range of transient phenomena 
(e.g., flares, radio bursts, coronal mass ejections) at many different time-scales, 
length-scales, and 
wavelengths. Observations and metadata concerning these phenomena are collected 
in the Heliophysics Event Knowledgebase (HEK) \citep{hurlburt2012}.  Entries are generated both by
automated algorithms and human observers.  Some of the information in the HEK 
reproduces feature and event data from elsewhere (for example, the \textit{GOES} flare catalogue),
and some is generated by the Solar Dynamics Observatory Feature Finding Team 
\citep{martens2012}.  A key feature of the HEK is that it
provides an homogeneous and well-described interface to a large amount of 
feature and event information. SunPy 
accesses this information through the \texttt{hek} module.  The \texttt{hek} module makes use of the 
HEK public API\footnote{For more information see \url{http://vso.stanford.edu/hekwiki/ApplicationProgrammingInterface}}.

@@

HEK comparisons can be combined using Python's logical operators (e.g., \texttt{and}
and \texttt{or}). The ability to use comparison and logical operators on HEK attributes allows 
the construction of queries of arbitrary complexity.
For the query in Listing \ref{code:hek:or} returns
returns flares with helioprojective x coordinates west of 50 arcseconds or 
those that have a peak flux above 1000.0 in units defined by the FRM.

@@

Frequently after users have found events of interest the next step is to 
download observational data. The \texttt{H2VClient} module makes this
east by providing a translation layer between HEK query results
and VSO data queries. This capability is demonstrated in Listing~\ref{code:hek2vso}.
\begin{listing}[H]
\pythoncode{pycode_hek4.txt}
\caption{Code snippet continuing from Listing~\ref{code:hek:or} showing the 
query and download of data from the first HEK result from the VSO.}
\label{code:hek2vso}
\end{listing}

### 2014-08-28 3-4-Helioviewer.tex
edits from full read thru

SunPy provides the ability to download images hosted by the
Helioviewer Project (\href{http://helioviewer.org}{http://helioviewer.org}).  
The aim of the Helioviewer Project is to enable
the exploration of solar and heliospheric data from multiple data
sources (such as instrumentation and feature/event catalogues) via
easy-to-use visual interfaces. The Helioviewer Project have developed two client 
applications that allow users to browse images and create movies of the Sun taken 
by a variety of instruments: \href{http://www.helioviewer.org}{Helioviewer.org}, a 
Google Maps-like web application, and \href{http://www.jhelioviewer.org}{JHelioviewer}, 
a movie streaming desktop application. In order to manage all of this data, the Helioviewer
project maintains archives of all of its data in JPEG2000 format (for further
details on the Helioviewer Project, see \citealt{muller2009}). The
JPEG2000 files are typically highly compressed compared to the source
FITS files from which they are generated, but are still high-fidelity, and thus can be used to quickly
visualise large amounts of data from multiple sources.  SunPy is
also used in Helioviewer production servers to manage the download and
ingestion of JPEG2000 files from remote servers.

@@

Reading JPEG2000 file into a SunPy session requires installing two
other pieces of software. The first, 
\href{http://www.openjpeg.org}{\texttt{OpenJPEG}}, is an open-source
library for reading and writing JPEG2000 files.  The other package 
required is 
\href{https://github.com/quintusdias/glymur}{\texttt{glymur}}, an
interface between Python and the OpenJPEG libraries (note that these
packages are {\it not} required to use the functionality described in
Section \ref{sssec:hv:png}).

>>>>

SunPy provides the ability to download images hosted by the
Helioviewer Project (\url{http://helioviewer.org}).  
The aim of the Helioviewer Project is to enable
the exploration of solar and heliospheric data from multiple data
sources (such as instrumentation and feature/event catalogues) via
easy-to-use visual interfaces. The Helioviewer Project have developed two client 
applications that allow users to browse images and create movies of the Sun taken 
by a variety of instruments: \url{http://www.helioviewer.org}, a 
Google Maps-like web application, and \url{http://www.jhelioviewer.org}, 
a movie streaming desktop application. In order to manage all of this data, the Helioviewer
project maintains archives of all of its data in JPEG2000 format (for further
details on the Helioviewer Project, see \citealt{muller2009}). The
JPEG2000 files are typically highly compressed compared to the source
FITS files from which they are generated, but are still high-fidelity, and thus can be used to quickly
visualise large amounts of data from multiple sources.  SunPy is
also used in Helioviewer production servers to manage the download and
ingestion of JPEG2000 files from remote servers.

@@

Reading JPEG2000 file into a SunPy session requires installing two
other pieces of software. The first, 
\url{http://www.openjpeg.org}{\texttt{OpenJPEG}}, is an open-source
library for reading and writing JPEG2000 files.  The other package 
required is 
\url{https://github.com/quintusdias/glymur}{\texttt{glymur}}, an
interface between Python and the OpenJPEG libraries (note that these
packages are {\it not} required to use the functionality described in
Section \ref{sssec:hv:png}).

### 2014-08-28 5-Dev.tex
edits from full read thru

>>>>

In order to enable the long-term development of SunPy, a formal organizational
structure has been defined. The management of SunPy is the responsibility of 
the SunPy board, a group of elected members of the community. The board elects
a lead developer whose is responsible for the day to day development of SunPy.
SunPy also makes use of Python-style Enhancement proposals which can be proposed
by the community and are voted on by the board. These proposals set the overal
direction of SunPy's development.
